Quantum Mechanics Assignment Three

### 2016-02-02 quantum/latex/a4.tex
all
Ssstuffff

\section{10.8: Magnetic Perturbation Spin 1}
For this problem we want to find the energy perturbations due to a magnetic field perturbation
in the x direction, while remaining in the z basis. To do this we first find the
perturbed Hamiltonian, which is just the x spin-1 matrix times a factor, and
then use the formula
%$E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}$
to find the second-order energy corrections.\\

@@

\begin{align*}
  E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}\\
\end{align*}

\begin{align*}
  E_{\frac{-1}{2}} &=
  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
  = \frac{\hbar\omega_2^2}{2\omega_0}\\
\end{align*}

\begin{align*}
  E_{0} &=
  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
  = \frac{\hbar\omega_2^2}{2\omega_0}\\
\end{align*}

\begin{align*}
  E_{\frac{1}{2}} &=
  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
  = \frac{\hbar\omega_2^2}{2\omega_0}\\
\end{align*}

>>>>

\section{10.8: Magnetic Perturbation Spin 1}
For this problem we want to find the energy perturbations due to a magnetic field perturbation
in the x direction, while remaining in the z basis. To do this we first find the
perturbed Hamiltonian, which is just the x spin-1 matrix times a factor, and
then use the formula
$E_n^{(2)} = \sum_{m\neq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}}|^2}{(E_n^{(0)}-E_m^{(0)})}$
to find the second-order energy corrections.\\

@@

\begin{align*}
  E_n^{(2)} = \sum_{m\neq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}}{(E_n^{(0)}-E_m^{(0)})}
\end{align*}

\begin{align*}
  E_1^{(2)} &=
  \frac{|H'_{1,2}|^2}{E_2 - E_1} +
  \frac{|H'_{1,3}|^2}{E_{3} - E_{1}}\\
  &= \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
  &= \frac{\hbar\omega_2^2}{4\omega_0}\\
\end{align*}

\begin{align*}
  E_2^{(2)} &=
  \frac{|H'_{2,1}|^2}{E_1 - E_2} +
  \frac{|H'_{2,3}|^2}{E_3 - E_2}\\
  &= -\frac{\hbar^2\omega_2^2}{4\hbar\omega_0} + \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
  &= 0\\
\end{align*}

\begin{align*}
  E_3^{(2)} &=
  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
  \frac{|H'_{3,2}|^2}{E_2 - E_3}\\
  &= -\frac{\hbar\omega_2^2}{4\omega_0}\\
\end{align*}

These are nicely symmetric energy corrections typical of physics problems. From this we can see
that the ratio between the initial and perturbation Larmor frequencies is the only factor in
deciding the magnitude of the perturbation. It is kind of weird that having a big initial
Larmor frequency decreases the magnitude of the perturbation energy.\\

We now plot our results:\\
\vspace{7cm}

As expected, we see quadratic dependence of the energy change on on perturbation strength for the
second-order correction.
\section{9.17: Ramp correction to Infinite Square Well}
Here we find the first-order energy corrections for an infinite square well wavefunction where a
ramp potential $H' = V'(x) = \beta x$ is applied. Because the correction Hamiltonian is in the
x basis, we'll need to use wave function notation to find the ground-state energy correction:

\begin{align*}
  E_1^{(1)} &= \bra{1}H'\ket{1}\\
  &= \frac{2\beta}{L}\int_0^L \sin^2\left(\frac{\pi x}{L}\right)x dx\\
  &= \frac{\beta L}{2}\\
\end{align*}

This shows how easy it is to apply perturbation theory to find the energy corrections for simple
systems. The units make sense and as we expect, both $\beta$ and L increase the energy shift.\\

\section{10.18: Dirac delta Infinite Square Well}
In this problem we deal with an infinite square well which has been perturbed by a delta function:
$H' = V(x) = LV_0\delta(x-L/2)$\\

\textbf{a.} First, we calculate the nth energy perturbation $E_n^1$ using wave function notation:

\begin{align*}
  E_n^{(1)} &= \bra{n}H'\ket{n}\\
  &= \int_0^L 2\sin^2\left(\frac{n\pi x}{L}\right)V_0\delta(x-L/2)dx\\
  &= 2V_0\sin^2\left(\frac{n\pi}{2}\right)\\
  \begin{cases}
    2V_0 & \mbox{n is odd}\\
    0 & \mbox{n is even}\\
  \end{cases}\\
\end{align*}

\textbf{b.}
This result is based on the parity of the function. Even-parity functions have zero density at the
center of the well, whereas odd-parity functions have nonzero density there. Therefore conceptually
we can say that the delta potential only interacts with the odd functions; the even functions tunnel
right through.\\

\textbf{c.}
The corrected wave function is given by:

\begin{align*}
  \ket{1} = \sum_{m\neq1} \frac{2mL^2\bra{m}H'\ket{m}}{\left(m^2-1\right)\hbar^2\pi^2}\\
\end{align*}

This value will be maximized when m is the smallest odd number larger than 1, i.e, 3. Thus the
ground state is corrected most by the $\ket{3^{(0)}}$ state. This makes sense, since in general
similar-energy waves correct each other the most. This conceptually makes sense, since the wave
would take on characteristics of those similar to it in energy.\\
